Why do we Eval?\\\Large A large-scale study of Eval usage in R

### 2021-03-27 paper/main.tex
Correct Kaggle numbers

\paragraph{Kaggle} In the Kaggle dataset, \kaggleNbruns packages trigger at least one \eval. In total, \kaggleAllcalls \eval were recorded. Most of them are calls to the normal \eval, whereas a few, \kaggleLocals, are calls to \c{local}. This happens in the machinery that loads library code.
Out of \kaggleStaticeval static \eval sites, we exercise \kaggleTriggeredeval of them. Mot \eval are invoked only once. Only one \eval amounts for several thousand, \kaggleMaxcalls,  invocations. Code amounts for \kaggleCodepercent of arguments in the kaggle corpus. However, contrary to the CRAN corpus, they are all expressions, and there are very few symbols (\kaggleSymbolpercent) and no language objects. Most expression objects result from a call to the \c{parse} function, suggesting that most \eval in Kaggle parse and evaluate strings. The median argument size is \kaggleMedianszeval, which makes sense, as there are very few arguments that are only symbols. The largest argument size is \kaggleMaxszeval. The distribution of events is similar to the one for CRAN or \emph{base}: \kaggleSmalleventspct of \eval calls amount for less than 50 events.

@@

\subsection{Normalized expressions} During tracing, we gather all source expressions and all resolved expression arguments passed to \eval. For instance, a non-resolved expression and some of its resolved expressions can be:

@@

Note that a site can exhibit several minimized expressions. However, \packageNbOneMinimizedPercent of the sites have only one minimized expression. Note also that if a function is called in that expression, we see only the function call but not the body of the function. We will trace what happens during the evaluation of the expression by \eval in Section~\ref{sec:dangerseval}, especially about side effects performed in \eval.

\paragraph{Variables.}
The most common minimized expression is a variable.\footnote{An arithmetic expression with some variables and constant will also be normalized to \c{X}, i.e. a variable.} It happens in \packageMinimizedpropsitesa of the sites and is widespread across packages as it is found in \packageMinimizedpackagea packages across our CRAN corpus. The \c{envir} argument is explicitly passed in \packageMinimizedpercentenvira\% of the call sites, which suggests that \eval is often used here to get the value of the variable in a non-local environment. If we look only at the \emph{symbols} passed to \eval, the percentage of non-default environments goes up to \packageNbNonDefaultEnvirVariablesPercent.

% the percentage of non-default environment is more convincing when looking at the number of calls rather the number of sites

@@

% for this one, currently uses any expression with $, [] or [[]] but we should rather use a more restrictive way (take only the ones without interesting calls)
Minimized expression \texttt{\$} refers to a slot access as the top call in a simple expression, for instance \texttt{a\$slot + 1}. It amounts for \packageGeneralizedSlotAccess sites, \ie \packageGeneralizedSlotAccessPercent of the sites.

@@

For the first category, \eval is not necessary. In \packageNbCallSitesUniqueActualValue sites exhibiting \c{V}, \ie \packageCallSitesUniqueActualValuePercent\% sites, it is only passed such actual values. The median number of runs for those sites is \packageMedianRunSitesUniqueActualValue (average \packageAverageRunSitesUniqueActualValue, standard deviation \packageSdRunSitesUniqueActualValue), suggesting it happens because there are rarely triggered in our dataset and all paths are not taken. However, manual inspection of the source code of some examples shows that some sites actually pass values. For instance, in package \emph{ctmle}, in function \c{evaluate\_candidate\_glmnet}:

@@

\paragraph{\c{model.frame}} This is a function used to create statistical models. We explain it in more details in the section about pattern (see Section~REF).

@@

The most common combinations for sites with several minimized expressions are shown in Figure~\ref{fig:frequentcombinationminimized}.

\begin{figure}
	\centering 
	\includegraphics[width=\textwidth]{package_combination_minimized.pdf}
	\caption{Most frequent combinations of minimized expressions per site, with a cumulative share of 95\% of the sites.} \label{fig:frequentcombinationminimized}
\end{figure}

@@

\NOTE{Two small paragraphs about normalized expressions in Base and Kaggle.}

@@

\mypara{Scope.} The environment argument can be \c{NULL}, a \c{List}
or a data frame. This happens in 3.5\% of the cases: \eval copies the
fields of the list or data frame and creates bindings for them in a
new environment. This pattern is used to evaluate formulas which can
directly refer to the fields of the data. The \c{envir} argument can
also be a number $n$. It means that the environment in which the
expressions is evaluated will be the result of \c{sys.call(n)} where
$n$ refers to the $n$-th stack frame.

\section{\eval patterns}

>>>>

\paragraph{Kaggle} In the Kaggle dataset, \kaggleNbruns notebooks trigger at least one \eval. In total, \kaggleAllcalls \eval were recorded. All of them are calls to the normal \eval.
Out of \kaggleStaticeval static \eval sites, we exercise \kaggleTriggeredeval of them. Most \eval are invoked only once. Only one \eval amounts for several thousand, \kaggleMaxcalls,  invocations. Code amounts for \kaggleCodepercent of arguments in the kaggle corpus. However, contrary to the CRAN corpus, they are all expressions, and there are very few symbols (\kaggleSymbolpercent) and no language objects. Most expression objects result from a call to the \c{parse} function, suggesting that most \eval in Kaggle parse and evaluate strings. The median argument size is \kaggleMedianszeval, which makes sense, as there are very few arguments that are only symbols. The largest argument size is \kaggleMaxszeval. The distribution of events is similar to the one for CRAN or \emph{base}: \kaggleSmalleventspct of \eval calls amount for less than 50 events.

@@

\subsection{Arguments of \eval}

\eval has three arguments, two of which we study in more details:

\begin{itemize}
	\item \c{expr}: the expression to evaluate
	\item \c{envir}: the environment in which performing the evaluation
	\item \c{enclos}
\end{itemize}

\subsection{The environment passed to \eval.}

The environment argument can be \c{NULL}, a \c{List}
or a data frame. This happens in 3.5\% of the cases: \eval copies the
fields of the list or data frame and creates bindings for them in a
new environment. This pattern is used to evaluate formulas which can
directly refer to the fields of the data. The \c{envir} argument can
also be a number $n$. It means that the environment in which the
expressions is evaluated will be the result of \c{sys.call(n)} where
$n$ refers to the $n$-th stack frame.

\subsubsection{The expression passed to \eval.}

During tracing, we gather all source expressions and all resolved expression arguments passed to \eval. For instance, a non-resolved expression and some of its resolved expressions can be:

@@

Note that a site can exhibit several minimized expressions (see Figure~\ref{fig:frequentcombinationminimized}). However, \packageNbOneMinimizedPercent of the sites have only one minimized expression. Note also that if a function is called in that expression, we see only the function call but not the body of the function. We will trace what happens during the evaluation of the expression by \eval in Section~\ref{sec:dangerseval}, especially about side effects performed in \eval.

\begin{figure}[htbp]
	\centering 
	\includegraphics[width=\textwidth]{package_combination_minimized.pdf}
	\caption{Most frequent combinations of minimized expressions per site, with a cumulative share of 95\% of the sites.} \label{fig:frequentcombinationminimized}
\end{figure}

\paragraph{Variables.}
The most common minimized expression is a variable.\footnote{An arithmetic expression with some variables and constant, \eg \c{1 + x}, will also be normalized to \c{X}, i.e. a variable.} It happens in \packageMinimizedpropsitesa\% of the sites and is widespread across packages as it is found in \packageMinimizedpackagea packages across our CRAN corpus. \packageNbSymbolVarSitePercent are only symbols, \ie not complex expressions. The \c{envir} argument is explicitly passed in \packageMinimizedpercentenvira\% of the call sites. The call sites without an explicit \c{envir} argument all use a specific \eval variant, \c{eval.parent}.

@@

Minimized expression \texttt{\$} refers to a slot access, with \texttt{\$}, \texttt{[]} or \texttt{[[]]}, as the top call in a simple expression, for instance \texttt{a\$slot + 1} and  It amounts for \packageNbSlotAccess sites, \ie \packageSlotAccessPercent of the sites.

As a whole, variable access, either as the top call, or inside a function call, possibly with accessing a slot, represents \packageGeneralizedVarAccessSitePercent of the sites.

@@

For the first category, \eval is not necessary. In \packageNbCallSitesUniqueActualValue sites exhibiting \c{V}, \ie \packageCallSitesUniqueActualValuePercent\% sites, it is only passed such actual values. The median number of runs for those sites is \packageMedianRunSitesUniqueActualValue (average \packageAverageRunSitesUniqueActualValue, standard deviation \packageSdRunSitesUniqueActualValue), suggesting it happens because of a lack of coverage, because there are rarely triggered in our dataset and all paths are not taken. However, manual inspection of the source code of some examples shows that some sites actually pass values. For instance, in package \emph{ctmle}, in function \c{evaluate\_candidate\_glmnet}:

@@

\paragraph{\c{model.frame}} This is a function used to create statistical models. We explain it in more details in the section about the usage of \eval (see Section~REF).

@@

@@

\paragraph{Base.} The ranking of minimized expressions in \emph{base} is similar to the one in CRAN bu the share of \c{V} is much larger, as it is present in \baseMinimizedpropsitesa\% of the sites and \c{X} is a distant second, in \baseMinimizedpropsitesb\% of the sites.

\paragraph{Kaggle.} Kaggle is dominated by sites with minimized expression \c{V}, with \kaggleMinimizedcallsitesa sites, and  \kaggleMinimizedpropsitesa\% of the \kaggleNbCallSites sites in total. All of those values are an AST size of one node, suggesting they do not come. \NOTE{TODO}
%TODO: all the src ref here seem to be incorrect and not from kaggle.

@@

\section{\eval patterns and anti-patterns}
